\begin{abstract}
Logical reasoning systems need different kinds of evidence for true, false, and unsupported queries, but most language-model pipelines still emphasize successful proofs and give much less guidance on refutation and abstention. We introduce \model, a supervision format with three verifiable response types: \prove{} for proving the query, \refute{} for deriving the opposite literal, and \abstain{} for emitting a failed-rule or missing-support witness. We pair this format with a symbolic verifier that checks whether generated traces are consistent with the theory closure. Our focus is not formal theorem proving or counterexample search in Lean. Instead, we study natural-language rule reasoning with true/false/unknown queries, where unsupported cases require checkable explanations of missing support. On a fixed three-seed 7B suite with only 4,096 training questions, \model{} does not maximize raw accuracy, but it consistently improves verifier-backed reasoning: joint label-plus-evidence accuracy rises from 39.9\% to 75.0\% in-domain and from 26.5\% to 55.0\% on depth-5, while raw accuracy remains close to proof-only supervision and within 1.5--3.6 points of answer-only. The main effect is better abstention, not simply more abstention: on gold-unknown queries, proof-only and \model{} predict \texttt{Unknown} at similar rates, but \model{} produces 1,444 faithful unknown explanations in-domain versus 0 for proof-only supervision. As training scale increases, the raw-accuracy gap largely disappears. At full train, \model{} reaches 98.2\% in-domain accuracy with 96.8\% joint accuracy, and on depth-5 it attains 93.8\% accuracy and 86.2\% joint accuracy on the fixed subset, slightly surpassing answer-only in raw accuracy. These results support a simple claim: supervising verifiable abstention witnesses is an effective way to improve faithful logical reasoning, and at sufficient scale it can deliver large faithfulness gains without sacrificing strong answer accuracy.
\end{abstract}
