\section{Discussion and Conclusion}
\label{sec:discussion}

\subsection{What the Full Study Supports}

The completed study supports a stronger and narrower claim than the pilot alone. At 0.5B, tri-modal supervision mainly improves faithfulness. At 7B, the core result is not universal raw-accuracy dominance, but a robust gain in verifier-backed reasoning quality. On the fixed 4k seeded suite, \model{} more than doubles proof-only joint accuracy in-domain and on depth-OOD. As the training budget grows, that reasoning gain persists while the raw-accuracy gap steadily contracts. In other words, the tradeoff changes with scale rather than reflecting a fixed weakness of the method.

\subsection{Why \model{} Helps}

The most important insight is that the benefit does not come primarily from predicting \texttt{Unknown} more often. At 7B, proof-only and \model{} abstain on nearly the same number of gold-unknown examples. The difference is that proof-only abstentions remain generic and unverifiable, whereas \model{} abstentions become explicit missing-support witnesses. The key improvement is therefore not the abstention label by itself, but the explanation attached to it. The abstention ablations reinforce this interpretation: both failure tags and witness text matter, and the full target works best when the two are combined. A second ablation adds an important qualifier. When we remove explicit refutation evidence on false cases, much of the faithful-unknown behavior remains, but overall faithfulness and joint accuracy drop substantially. The clearest unique gain is therefore on unsupported queries, while full verifier-backed performance still benefits from explicit refutation supervision.

\subsection{Relation to Concurrent Formal Disproof Work}

Concurrent work by \citet{li2026learningtodisprove} studies formal counterexample generation in Lean using symbolic mutation and multi-reward expert iteration. That paper strengthens the broader motivation for moving beyond proof-only supervision, but it also makes our scope clearer. Their contribution is a data-generation and training framework for machine-checked counterexamples to false formal statements. Ours is a supervision and evaluation study for natural-language logical reasoning with three kinds of queries, including unsupported queries that require abstention witnesses. The overlap is in motivation, not in method.

\subsection{Scaling Behavior}

Scaling changes how the paper should be read. At low budget, \model{} buys much better faithfulness at a modest raw-accuracy cost. At larger budgets, that summary becomes incomplete. On the full-train fixed depth-5 subset, \model{} slightly surpasses answer-only in raw accuracy, and on several transfer settings it remains competitive while preserving very large joint-accuracy gains. The practical lesson is that tri-modal supervision becomes more attractive, not less, as the base model and training budget improve.

\subsection{Limitations}

The paper is still not a ``solved logical reasoning'' result. First, the study uses one benchmark family and one main 7B backbone, so the strongest claims should be restricted to ProofWriter-style reasoning. Second, answer-only systems receive joint accuracy 0 by construction because they do not emit evidence; this is appropriate for a faithfulness paper, but it means that joint accuracy should be interpreted alongside raw accuracy rather than in isolation. Third, the ablation studies are single-seed measurements, which are enough to localize the effect but not enough to fully characterize variance.

\subsection{Next Steps}

The natural next step is to stress-test the same supervision idea beyond the current setting. The most useful directions are:
\begin{itemize}
\item \textbf{backbone breadth}: repeat the same supervision comparison on additional 7B--14B models to test whether the scaling pattern is architecture-specific,
\item \textbf{benchmark breadth}: evaluate on other verifier-backed reasoning datasets where abstention and refutation matter,
\item \textbf{harder OOD settings}: target the remaining gap on long refutation chains and more adversarial distribution shifts,
\item \textbf{answer/rationale decoupling}: combine a short answer head with a verifier-constrained rationale channel so that raw classification and evidence generation can specialize.
\end{itemize}

\subsection{Conclusion}

We presented \model, a tri-modal distillation format for logical reasoning. The main result is not that proof/refutation/abstention supervision dominates every raw metric, but that it reliably improves verifier-backed reasoning and abstention faithfulness, and that this advantage survives seeds, scaling, and transfer. Proof-only supervision can learn to output structured traces, but it still fails to make abstention evidence verifiable. \model{} closes that gap, making verifier-backed abstention and refutation practical supervision targets for faithful logical reasoning while leaving broader benchmark transfer and harder long-depth reasoning as the next frontier.
