\section{Appendix}

\subsection{Prompt Formats}

All variants use the same fact and rule context and differ only in the target format. The prompt prefix is:
\begin{verbatim}
You are a careful logical reasoner.
Use only the numbered facts and rules below.
\end{verbatim}

The main task formats are:
\begin{itemize}
\item \textbf{answer-only}: \texttt{LABEL: <True|False|Unknown>}
\item \textbf{proof-only}: \texttt{MODE}, \texttt{LABEL}, \texttt{CHAIN}, and \texttt{WITNESS}, with \texttt{CHAIN} allowed to be \texttt{NONE}
\item \model{} and its ablations: the same four fields, with \texttt{CHAIN} used for either a derivation path or a \texttt{FAIL[...]} abstention tag
\end{itemize}

This shared prompt structure matters for the comparison. The main difference between variants is the supervision target, not the surrounding instruction.

\subsection{Training Hyperparameters}

All runs use QLoRA with:
\begin{itemize}
\item 4-bit NF4 loading,
\item LoRA rank 16,
\item LoRA alpha 32,
\item dropout 0.05,
\item max sequence length 512,
\item learning rate \(2\times 10^{-4}\),
\item one training epoch.
\end{itemize}

For 7B runs, the per-device train batch size is 2 with gradient accumulation 8.
The seeded 4k suite changes the sampled training questions across seeds while
holding the evaluation subset fixed. Experiments were run on NVIDIA A100 80GB
GPUs.

\subsection{Verifier Details}

The verifier replays generated outputs against the benchmark's symbolic structure:
\begin{itemize}
\item For \textbf{PROVE} outputs, the predicted \texttt{CHAIN} must derive the queried literal.
\item For \textbf{REFUTE} outputs, the predicted \texttt{CHAIN} must derive the opposite literal.
\item For \textbf{ABSTAIN} outputs, the \texttt{WITNESS} must describe missing support from the canonical blocked derivation, or the chain must match the canonical \texttt{FAIL[no\_rule]} target for that query.
\end{itemize}

This verifier is deliberately lightweight. It does not search for an alternative proof on the model's behalf. It only checks the consistency of the model's own evidence claim. The target constructor, however, does run a closure computation before any \texttt{FAIL[no\_rule]} target is created. Thus a bare \texttt{FAIL[no\_rule]} label is not accepted for arbitrary unknown outputs; it receives credit only when the query's canonical blocked-derivation target is also \texttt{FAIL[no\_rule]}.

\paragraph{Abstention target construction.}
The target constructor uses the following logic, with a maximum recursive
failure depth of 3 in our experiments:
\begin{verbatim}
closure = forward_chain(facts, rules)
if query in closure: emit PROVE target
if opposite(query) in closure: emit REFUTE target

def explain(target, depth):
    if target in closure:
        return PROVED
    candidates = []
    for rule in rules:
        for binding in unify(rule.consequent, target):
            missing = []
            for antecedent in rule.antecedents:
                grounded = instantiate(antecedent, binding)
                if grounded not in closure:
                    missing.append(explain(grounded, depth - 1)
                                   if depth > 0 else MISSING(grounded))
            if missing:
                candidates.append(FAIL_RULE(rule.id, missing))
    if candidates is empty:
        return FAIL_NO_RULE(target)
    return argmin(candidates,
                  key=(number_of_missing_leaves,
                       failure_depth,
                       rule_id))
\end{verbatim}

This procedure makes the path-relevance assumption explicit: a
\texttt{FAIL[ruleX]} witness must correspond to a rule whose consequent can
derive the current target under some grounding, and its missing leaves are the
antecedents that block that grounded rule application. It does not enumerate all
minimal hitting sets of all possible derivations.

\paragraph{Abstention verification.}
Given a parsed model output for a gold-\texttt{Unknown} query, the main verifier
uses:
\begin{verbatim}
canonical = explain(query, depth=3)
if mode != ABSTAIN:
    return false
if normalized(chain) == canonical.tag:
    return true
if normalized(witness) == canonical.witness:
    return true
if every canonical missing leaf is mentioned in witness:
    return true
return false
\end{verbatim}
The strict audit in Table~\ref{tab:abstention-audit} replaces the three
acceptance routes with a single exact condition:
\begin{verbatim}
normalized(chain) == canonical.tag
and normalized(witness) == canonical.witness
\end{verbatim}
This stricter metric is closer to target imitation and penalizes valid but
non-canonical blocked derivations. The main metric is closer to tolerant
semantic checking, but does not prove minimality or uniqueness.

\subsection{Experiment Automation}

Main-track sweeps are launched with a single manifest-driven script:
\begin{verbatim}
python scripts/run_main_track_suite.py
\end{verbatim}
This launcher keeps per-run metadata, GPU-specific logs, and deterministic output names. After new result JSONs are produced, summaries, plots, LaTeX tables, and unknown-behavior reports are refreshed with:
\begin{verbatim}
python scripts/refresh_artifacts.py --compile-paper
\end{verbatim}

\subsection{Claim-to-Evidence Map}

Table~\ref{tab:claims-map} links each headline claim in the paper to one exact result artifact and states the scope of that claim. We use this table as a guardrail against vague summary language.

\begin{table}[t]
\centering
\small
\input{generated/claims_table.tex}
\caption{Claim-to-evidence map for the main paper. Each row states one headline claim, the exact number or table that supports it, and the scope of that evidence.}
\label{tab:claims-map}
\end{table}

\subsection{Additional Results}
\label{app:additional-results}

\begin{table}[t]
\centering
\small
\input{generated/backbone_table.tex}
\caption{Additional backbone check at 4k training examples. Qwen2.5-7B rows use
three seeds; Mistral-7B rows are a single seed.}
\label{tab:backbone-check}
\end{table}

\begin{table}[t]
\centering
\small
\input{generated/support_deletion_breakdown_table.tex}
\caption{Qwen2.5-7B support-deletion results by deleted support type.}
\label{tab:support-deletion-breakdown}
\end{table}

\begin{table}[t]
\centering
\small
\input{generated/ablation_table.tex}
\caption{Ablations on the abstention target.}
\label{tab:ablation-results}
\end{table}

\begin{table}[t]
\centering
\small
\input{generated/refute_ablation_table.tex}
\caption{Single-seed 4k ablation of explicit refutation evidence.}
\label{tab:refute-ablation-results}
\end{table}

\begin{table}[t]
\centering
\small
\input{generated/transfer_table.tex}
\caption{Transfer across additional ProofWriter test configurations.}
\label{tab:transfer-results}
\end{table}

These additional results support the same reading as the main experiments. The
Mistral check preserves the joint-accuracy gain, though it is only a single
seed. The ablations show that failed-rule tags and witness text each carry
signal, while explicit refutation evidence is still needed for full
label-plus-verifier-accepted-evidence coverage.

\subsection{Per-Class Breakdown on the 7B Seeded Suite}

Table~\ref{tab:per-class-seeded} reports per-class F1 on the fixed 4k seeded 7B suite. At this training budget, answer-only remains the strongest raw classifier by classwise F1. The main advantage of \model{} at this scale appears instead in verifier-accepted evidence and joint accuracy.

\begin{table}[t]
\centering
\small
\input{generated/per_class_table.tex}
\caption{Per-class F1 on the fixed 7B seeded suite. This table helps separate raw class prediction from verifier-accepted evidence.}
\label{tab:per-class-seeded}
\end{table}

This table also checks for side effects on supported queries at the small 4k
budget. In-domain, \model{} is within 1.3 F1 points of proof-only on True and
within 0.2 on False. On depth-OOD, \model{} is within 0.3 F1 points on True and
0.2 on False. The main residual cost is therefore not a large collapse on
PROVE/REFUTE quality, but the broader raw-accuracy gap to answer-only reported
in Table~\ref{tab:seeded-results}.

\subsection{Confusion Matrix on the Fixed 7B Suite}

Table~\ref{tab:confusion-matrix} expands the error profile into a full
gold-versus-predicted label matrix. It makes the same point as the main-text
error-profile table: the main added conservatism of \model{} shows up as a
moderate increase in predicted \texttt{Unknown} on supported queries, not as a
collapse in supported True/False classification.

\begin{table}[t]
\centering
\small
\input{generated/confusion_matrix_table.tex}
\caption{Confusion matrix on the fixed 7B seeded suite. The rows are grouped by
gold label and reported separately for each variant.}
\label{tab:confusion-matrix}
\end{table}

\subsection{Metric Interpretation}

Joint accuracy is defined as correct label plus verifier-accepted evidence. This
gives answer-only joint accuracy 0 by construction, because answer-only systems
do not emit evidence to verify. We therefore read joint accuracy together with
raw accuracy throughout the paper: answer-only measures label prediction alone,
while proof-only and \model{} are asked to produce checkable reasoning.
Verifier-accepted unknown evidence is the fraction of gold-\texttt{Unknown}
examples for which the model predicts \texttt{Unknown} and emits an accepted
abstention witness. Strict unknown evidence is the fraction of
gold-\texttt{Unknown} examples for which the predicted tag and witness exactly
match the canonical target.

\subsection{Post-hoc Symbolic Repair Baseline}

The main comparison asks whether supervision alone can make the model emit
checkable abstention evidence. A stronger solver-assisted alternative is to
attach the canonical witness from the same closure search whenever the model
already predicts \texttt{Unknown}. Table~\ref{tab:symbolic-repair} reports the
solver-assisted baseline for the 7B seeded suite. It is useful context, but it
is not model-generated evidence.

\begin{table}[t]
\centering
\small
\input{generated/symbolic_repair_table.tex}
\caption{Solver-assisted repair under fixed model labels on the fixed 7B seeded
suite. This is not model-generated evidence: the repaired column preserves the
model's predicted labels, attaches canonical witnesses only after the model
predicts \texttt{Unknown}, and therefore contextualizes solver-assisted
test-time repair rather than replacing the main supervision-only comparison.
Rows report 3-seed means.}
\label{tab:symbolic-repair}
\end{table}

\subsection{Qualitative Examples and Failure Modes}

The following examples are drawn from the fixed in-domain 4k Qwen2.5-7B suite.
They illustrate one canonical \texttt{FAIL[no\_rule]} success, one
\texttt{FAIL[ruleX]} case that tolerant verification accepts because the
canonical missing leaves are present, one proof-only generic abstention, and
one \texttt{FAIL[ruleX]} case where strict canonical matching fails.

\input{generated/qualitative_examples.tex}

\subsection{Claim Audit Artifact}

Before any submission draft leaves the repository, we run:
\begin{verbatim}
python scripts/check_paper_claims.py
\end{verbatim}
This script checks the headline numbers in the abstract, introduction, and discussion against the aggregated result files and writes the audit to \texttt{docs/PAPER\_CLAIM\_CHECK\_2026-04-26.md}. The current audit passes.
